---
title: "ORA of clustering results"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BuildNetwork}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, cache=TRUE,
  comment = "#>"
)
```

```{r setup}
library(AnNet)
library(synaptome.db)
library(pander)
library(ggplot2)
library(rSpectral)
library(data.table)
library(fgsea)
scale <- function(x, VALUE=NULL){

    x = as.numeric(as.vector(x))

    xmin <- min(x,na.rm=T)
    xmax <- max(x,na.rm=T)
   
    if( is.null(VALUE) ){

        x  <- x-xmin
        x  <- ifelse(!is.na(x), x/(xmax-xmin), NA) 

        return(x)
    }

    value = as.numeric(as.vector(value)[1])
    value = value-xmin
    value = ifelse(!is.na(value), value/(xmax-xmin), NA) 
    return(value)
}
```
# Title   AnNet: library for comprehensive network analysis
##  Authors: Colin Mclean, Anatoly Sorokin, J. Douglas Armstrong and Oksana Sorokina    

## Get modularity. Normalised modularity.
Normalised modularity allows comparison the networks with different structure.
Qm based on the previous studies by Parter et al., 2007, Takemoto, 2012,
Takemoto, 2013, Takemoto and Borjigin, 2011, which was defined as:

$$Qm = \frac{Q_{real}-Q_{rand}}{Q_{max}-Q_{rand}}$$

Where $Q_{real}$ is the network modularity of a real-world signaling network and,
$Q_{rand}$ is the average network modularity value obtained from 10,000 randomized
networks constructed from its real-world network. 
$Q_{max}$ was estimated as: $$Q_{max}=1 − \frac{1}{M}$$, where $M$ is the number of 
modules in the real network.
Randomized networks were generated from a real-world network using the
edge-rewiring algorithm (Maslov and Sneppen, 2002).

#Introduction

```{r read.net}
file <- system.file("extdata", "PPI_Presynaptic.gml", package = "AnNet")
gg <- igraph::read.graph(file,format="gml")
summary(gg)
g<-gg
```
```
{r clean.algs}
alg=c('lec','wt','fc','infomap','louvain','sgG1','sgG2','sgG5','spectral')
g<-gg
for(a in alg){
  if(a %in% vertex_attr_names(g)){
    g<-delete_vertex_attr(g,a)
  }
}
summary(g)
```


# Clustering
```
{r cluster.graph, cache=TRUE}
g<-calcAllClustering(g)
```

# Centrality
```{r calc.centr}
g <- calcCentrality(g)
```

```{r make.annotations}
gda<-prepareGDA(g,'TopOntoOVGHDOID')
genes<-as.character(V(g))
anNames<-getAnnotationList(gda)
anL<-lapply(anNames, function(.a){genes[grepl(.a,gda,fixed = TRUE)]})
names(anL)<-unescapeAnnotation(anNames)
pander(sapply(anL,length))
```

```{r fora}
cl<-make_clusters(g,as.numeric(V(g)$louvain))
resL<-lapply(1:length(cl),function(.i){res<-fora(anL,as.character(V(g)[which(membership(cl)==.i)]),universe = as.character(V(g)));return(res[,cl:=.i])})
res<-do.call(rbind,resL)
res<-res[padj<0.1,]
res[,overlapGenes:=sapply(overlapGenes,paste,collapse = ', ')]
res<-res[,c("cl","pathway", "pval", "padj", "overlap", "size", "overlapGenes")]
pander(res)
```

```{r annotate.go}
sfile<-system.file("extdata", "flatfile.go.BP.csv", package = "AnNet")
goBP <- read.table(sfile,sep="\t",skip=1,header=F,strip.white=T,quote="")
g <- annotate_go_bp(g, goBP)
summary(g)

sfile<-system.file("extdata", "flatfile.go.MF.csv", package = "AnNet")
goMF <- read.table(sfile,sep="\t",skip=1,header=F,strip.white=T,quote="")
g <- annotate_go_mf(g, goMF)
summary(g)

sfile<-system.file("extdata", "flatfile.go.CC.csv", package = "AnNet")
goCC <- read.table(sfile,sep="\t",skip=1,header=F,strip.white=T,quote="")
g <- annotate_go_cc(g, goCC)
summary(g)

```

```{r short.fora}
anL<-getAnnotationVertexList(g,'GO_CC_ID')
res<-clusterORA(g,alg='louvain',name='GO_CC_ID',vid='name')
goCCdf<-unique(goCC[,1:2])
names(goCCdf)<-c('ID','Term')
rr<-merge(goCCdf,res,by.y='pathway',by.x='ID')
pander(rr[order(rr$cl),])
```

