---
title: "BuildNetwork"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BuildNetwork}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, cache=TRUE,
  comment = "#>"
)
```

```{r setup}
library(AnNet)
library(synaptome.db)
library(pander)
library(ggplot2)
library(topGO)
library(org.Hs.eg.db)
```
# Title   AnNet: library for comprehensive network analysis
##  Authors: Colin Mclean, Anatoly Sorokin, J. Douglas Armstrong and Oksana Sorokina    

#Introduction

# Overview of capabilities
# Build the network

The package allows building the network from the data frame, where the rows correspond to the edges of the graph, or for the list of nodes (genes), for which the information will be retrieved from the SynaptomeDB.

## Build network for the node list extracted from SynaptomeDB

```{r net_from_SynaptomeDB}

cid<-match('Presynaptic',getCompartments()$Name) # Let's get the ID for presynaptic compartment
cid
t<-getAllGenes4Compartment(cid) # Now we need to collect all the gene IDs for presinaptic  compartment
dim(t)
head(t)
gg<-buildFromSynaptomeByEntrez(t$HumanEntrez) # finally, build the graph using respecctive gene EntrezIDs as node IDs
summary(gg)

```
# Annotate the nodes with node attributes
## Gene name
```{r annotate_net}
gg<-annotateGeneNames(gg)
summary(gg)
head(V(gg))
head(V(gg)$GeneName)
```

# Prepare GO annotations

Define constants required for the analysis: specie (TAX), specie genomics data
base (TAXmap) and its name (TAXmapStr), minimum (thresMin) and maximum of 
genes annotated by the term to be included in the annotation list.
```{r get.go.init}
## The GO Ontology to run the enrichment
TAX       = "9606"
TAXmap    = org.Hs.eg.db
TAXmapStr = "org.Hs.eg.db"    
TAXtitle  = "Human"
thresMIN <- 10
thresMAX <- 1000
```

# CC ontology annotation

Set ontology name
```{r set.ont.cc}
Onto      = "CC"
```


```{r make.go.data.cc}
## Get Gene Symbols 
geneID2TERM <- revmap(topGO::annFUN.org(feasibleGenes=NULL,whichOnto=Onto, mapping = TAXmapStr, ID = "Entrez"))
TERM2geneID <- revmap(geneID2TERM)

geneIDs=names(geneID2TERM)

myList <- as.vector(V(gg)$name)
myList <- unique(myList)
myList <- na.omit(myList)
geneList <- factor(as.integer(geneIDs %in% myList))
names(geneList) <- geneIDs

## setup the topGO enrichment object 
GOdata <- new("topGOdata", ontology=Onto, allGenes = geneList, annot = annFUN.gene2GO, gene2GO = geneID2TERM)
```

```{r prepare.terms.cc}
##--- Run enrichment using topGO
resultFisher        <- runTest(GOdata, algorithm = "classic", statistic = "fisher")
resultFisher.elim   <- runTest(GOdata, algorithm = "elim", statistic = "fisher")
        
topnodes = resultFisher@geneData[[4]]

if( resultFisher.elim@geneData[[4]] < topnodes ){
    topnodes = resultFisher.elim@geneData[[4]]
}

## The GO object after running the enrichment analysis
allRes <- GenTable(GOdata, classicFisher = resultFisher, elimFisher = resultFisher.elim, orderBy = "classicFisher", ranksOf = "classicFisher", topNodes = topnodes )
##---

## Add Entrez IDs to output GO object
hits=lapply(allRes$GO.ID,function(x)
                          intersect(sigGenes(GOdata),genesInTerm(GOdata, c(x))[[1]]))
allRes$hits=hits
```

```{r prepare.annot.df.cc}
NGOterms = dim(allRes)[1]

flatfile <- data.frame()

for( i in 1:NGOterms ){

    ids  = allRes$hits[i][[1]]
    Nids = length(ids)
    if( Nids >= thresMIN && Nids <= thresMAX ){

        if( i == 1 ){
            flatfile = cbind(rep(allRes$GO.ID[i][[1]],Nids),
                             rep(allRes$Term[i][[1]], Nids),
                             ids)
        } else {
            tmp = cbind(rep(allRes$GO.ID[i][[1]],Nids),
                        rep(allRes$Term[i][[1]], Nids),
                        ids)
            flatfile = rbind(flatfile, tmp)
        }        

    }
    
}

```

Now we can annotate graph with new data:
```{r annot.graph.cc}
gg<-annotate_go_cc(gg,flatfile)
summary(gg)
```

You can see two new vertex attributes: O_CC_ID and GO_CC for lists of IDs and
terms respectively. 

For later use annotation data could be saved into the file:
```{r write.file.cc}
write.table(flatfile,
            sprintf("flatfile.go.%s.csv",Onto),
            sep="\t",
            col.names=F,
            row.names=F,
            quote=F)
```

# BP ontology annotation

Set ontology name
```{r set.ont.bp}
Onto      = "BP"
```


```{r make.go.data.bp}
## Get Gene Symbols 
geneID2TERM <- revmap(topGO::annFUN.org(feasibleGenes=NULL,whichOnto=Onto, mapping = TAXmapStr, ID = "Entrez"))
TERM2geneID <- revmap(geneID2TERM)

geneIDs=names(geneID2TERM)

myList <- as.vector(V(gg)$name)
myList <- unique(myList)
myList <- na.omit(myList)
geneList <- factor(as.integer(geneIDs %in% myList))
names(geneList) <- geneIDs

## setup the topGO enrichment object 
GOdata <- new("topGOdata", ontology=Onto, allGenes = geneList, annot = annFUN.gene2GO, gene2GO = geneID2TERM)
```

```{r prepare.terms.bp}
##--- Run enrichment using topGO
resultFisher        <- runTest(GOdata, algorithm = "classic", statistic = "fisher")
resultFisher.elim   <- runTest(GOdata, algorithm = "elim", statistic = "fisher")
        
topnodes = resultFisher@geneData[[4]]

if( resultFisher.elim@geneData[[4]] < topnodes ){
    topnodes = resultFisher.elim@geneData[[4]]
}

## The GO object after running the enrichment analysis
allRes <- GenTable(GOdata, classicFisher = resultFisher, elimFisher = resultFisher.elim, orderBy = "classicFisher", ranksOf = "classicFisher", topNodes = topnodes )
##---

## Add Entrez IDs to output GO object
hits=lapply(allRes$GO.ID,function(x)
                          intersect(sigGenes(GOdata),genesInTerm(GOdata, c(x))[[1]]))
allRes$hits=hits
```

```{r prepare.annot.df.bp}
NGOterms = dim(allRes)[1]

flatfile <- data.frame()

for( i in 1:NGOterms ){

    ids  = allRes$hits[i][[1]]
    Nids = length(ids)
    if( Nids >= thresMIN && Nids <= thresMAX ){

        if( i == 1 ){
            flatfile = cbind(rep(allRes$GO.ID[i][[1]],Nids),
                             rep(allRes$Term[i][[1]], Nids),
                             ids)
        } else {
            tmp = cbind(rep(allRes$GO.ID[i][[1]],Nids),
                        rep(allRes$Term[i][[1]], Nids),
                        ids)
            flatfile = rbind(flatfile, tmp)
        }        

    }
    
}

```

Now we can annotate graph with new data:
```{r annot.graph.bp}
gg<-annotate_go_bp(gg,flatfile)
summary(gg)
```

You can see two new vertex attributes: O_BP_ID and GO_BP for lists of IDs and
terms respectively. 

For later use annotation data could be saved into the file:
```{r write.file.bp}
write.table(flatfile,
            sprintf("flatfile.go.%s.csv",Onto),
            sep="\t",
            col.names=F,
            row.names=F,
            quote=F)
```


# MF ontology annotation

Set ontology name
```{r set.ont.mf}
Onto      = "MF"
```


```{r make.go.data.mf}
## Get Gene Symbols 
geneID2TERM <- revmap(topGO::annFUN.org(feasibleGenes=NULL,whichOnto=Onto, mapping = TAXmapStr, ID = "Entrez"))
TERM2geneID <- revmap(geneID2TERM)

geneIDs=names(geneID2TERM)

myList <- as.vector(V(gg)$name)
myList <- unique(myList)
myList <- na.omit(myList)
geneList <- factor(as.integer(geneIDs %in% myList))
names(geneList) <- geneIDs

## setup the topGO enrichment object 
GOdata <- new("topGOdata", ontology=Onto, allGenes = geneList, annot = annFUN.gene2GO, gene2GO = geneID2TERM)
```

```{r prepare.terms.mf}
##--- Run enrichment using topGO
resultFisher        <- runTest(GOdata, algorithm = "classic", statistic = "fisher")
resultFisher.elim   <- runTest(GOdata, algorithm = "elim", statistic = "fisher")
        
topnodes = resultFisher@geneData[[4]]

if( resultFisher.elim@geneData[[4]] < topnodes ){
    topnodes = resultFisher.elim@geneData[[4]]
}

## The GO object after running the enrichment analysis
allRes <- GenTable(GOdata, classicFisher = resultFisher, elimFisher = resultFisher.elim, orderBy = "classicFisher", ranksOf = "classicFisher", topNodes = topnodes )
##---

## Add Entrez IDs to output GO object
hits=lapply(allRes$GO.ID,function(x)
                          intersect(sigGenes(GOdata),genesInTerm(GOdata, c(x))[[1]]))
allRes$hits=hits
```

```{r prepare.annot.df.mf}
NGOterms = dim(allRes)[1]

flatfile <- data.frame()

for( i in 1:NGOterms ){

    ids  = allRes$hits[i][[1]]
    Nids = length(ids)
    if( Nids >= thresMIN && Nids <= thresMAX ){

        if( i == 1 ){
            flatfile = cbind(rep(allRes$GO.ID[i][[1]],Nids),
                             rep(allRes$Term[i][[1]], Nids),
                             ids)
        } else {
            tmp = cbind(rep(allRes$GO.ID[i][[1]],Nids),
                        rep(allRes$Term[i][[1]], Nids),
                        ids)
            flatfile = rbind(flatfile, tmp)
        }        

    }
    
}

```

Now we can annotate graph with new data:
```{r annot.graph.mf}
gg<-annotate_go_mf(gg,flatfile)
summary(gg)
```

You can see two new vertex attributes: O_MF_ID and GO_MFfor lists of IDs and
terms respectively. 

For later use annotation data could be saved into the file:
```{r write.file.mf}
write.table(flatfile,
            sprintf("flatfile.go.%s.csv",Onto),
            sep="\t",
            col.names=F,
            row.names=F,
            quote=F)
```
