---
title: "BuildNetwork"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BuildNetwork}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, cache=TRUE,
  comment = "#>"
)
```

```{r setup}
library(AnNet)
library(synaptome.db)
library(pander)
library(ggplot2)
library(topOnto) 
library(org.Hs.eg.db)
```
# Create ontology annotation for Disease Ontology with TopOnto
##  Authors: Colin Mclean, Anatoly Sorokin, J. Douglas Armstrong and Oksana Sorokina    

# Build the network

The package allows building the network from the data frame, where the rows correspond to the edges of the graph, or for the list of nodes (genes), for which the information will be retrieved from the SynaptomeDB.

## Build network for the node list extracted from SynaptomeDB

```{r net_from_SynaptomeDB}

cid<-match('Presynaptic',getCompartments()$Name) # Let's get the ID for presynaptic compartment
cid
t<-getAllGenes4Compartment(cid) # Now we need to collect all the gene IDs for presinaptic  compartment
dim(t)
head(t)
gg<-buildFromSynaptomeByEntrez(t$HumanEntrez) # finally, build the graph using respecctive gene EntrezIDs as node IDs
summary(gg)

```
# Annotate the nodes with node attributes
## Gene name
```{r annotate_net}
gg<-annotateGeneNames(gg)
summary(gg)
head(V(gg))
head(V(gg)$GeneName)
```

# Prepare Disease annotations

Extract disease annotations from synaptome.db:
```{r get.diseases}
myList <- as.vector(V(gg)$name)
myList <- unique(myList)
myList <- na.omit(myList)
t<-getGeneDiseaseByEntres(myList)
t<-t[,c('HDOID','Description','HumanEntrez','HumanName')]
```

and restrict the list to only subset of diseases:
```{r subset.diseases}
tD<-t[t$HDOID%in%getDiseases(),]
```

Now we can annotate graph with new data:
```{r annot.graph.mf}
gg<-annotate_go_mf(gg,flatfile)
summary(gg)
```

You can see two new vertex attributes: O_MF_ID and GO_MFfor lists of IDs and
terms respectively. 

For later use annotation data could be saved into the file:
```{r write.file.mf}
write.table(flatfile,
            sprintf("flatfile.go.%s.csv",Onto),
            sep="\t",
            col.names=F,
            row.names=F,
            quote=F)
```
