---
title: "Evaluation of effectiveness of the permutation test on centrality measures "
author: "A. Sorokin"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
output:
  pdf_document:
    keep_tex: yes
    number_sections: yes
  html_document: default
params:
  format: !r if(opts_knit$get("rmarkdown.pandoc.to") == 'html') c('screen', 'print')
    else 'print'
  version: !r if(nchar(Sys.which("git"))) system("git describe --long --dirty --abbrev=10  --tags  --always",
    intern=TRUE) else date()
header-includes:
- \usepackage[T2A]{fontenc}
- \usepackage[utf8]{inputenc}
- \usepackage[english,russian]{babel}
- \usepackage{grffile}
- \usepackage{rotating}
- \usepackage{caption}
- \usepackage{longtable}
- \usepackage{lscape}
---
```{r loadPackages, include=FALSE, cache=FALSE}
## load additional packages in this chunk
library(pander)
library(knitr)
library(ggplot2)
library(xtable)
library(data.table)
library(ggdendro)
library(tidyr)
library("FactoMineR")
library(dendextend)
library(factoextra)
library(Matrix)
library(wordspace)
library(AnNet)
library(RColorBrewer)
library(UpSetR)
library(ggvenn)
library(synaptome.db)
library(plyr)
```

```{r setup, include=FALSE, cache=FALSE}
## This chunk should contain global configuration commands.
## Use this to set knitr options and related things. Everything
## in this chunk will be included in an appendix to document the
## configuration used.
#output <- opts_knit$get("rmarkdown.pandoc.to")
opts_knit$set(stop_on_error = 2L)

## By default R code is only included in HTML versions of the report
## (where it can be collapsed). You can generate a PDF version
## using rmarkdown::pdf_document to get a copy for print. Extensive
## chunks of R code may or may not be desired in /hat setting. If you
## want them simply change the following arguments to `echo = TRUE`.
## In either case the default can be overwritten for individual chunks.
#opts_chunk$set(echo = output=="html")
#opts_chunk$set(warning = output=="html")
#opts_chunk$set(message = output=="html")

## Cache options
opts_chunk$set(cache=TRUE,warning=FALSE)

## Figure options
## Set default figure format
#options(reportmd.figure.format=params$format)

## Set 'hide.fig.code' to FALSE to include code chunks that
## produce Figures in the output. Note that this affects all chunks
## that provide a figure caption.
opts_chunk$set(hold=TRUE, hide.fig.code=FALSE)

## Set up default plotting options for different formats.
## These can be overwritten for individual chunks
#interactiveFig()
#screenFig()
#printFig()

## Pander options
panderOptions("digits", 3)
panderOptions("table.split.table", 160)
#panderOptions("table.style", "grid")

## Configure Figure and Table lables
#options(figcap.prefix = "Figure", figcap.sep = ":", figcap.prefix.highlight = "**")
#options(tabcap.prefix = "Table", tabcap.sep = ":", tabcap.prefix.highlight = "**")

## Install required knitr hooks
#installHooks()
```

```{r functions, include=FALSE}
## Custom functions used in the analysis should go into this chunk.
## They will be listed in their own section of the appendix.
getCM<-function(m,nm,keepOrder){
  v<-as.numeric(m[,which(colnames(m)==nm)])
  if(keepOrder){
    return(v)
  }else{
    return(sort(v,decreasing = FALSE,na.last=TRUE))
  }
}

calcCentralityInternalDistances<-function(l,keepOrder=FALSE,dist='euclidean'){
  CN  <- c("ID","DEG","BET","CC","SL","mnSP","PR","sdSP")
  resl<-list()
  for(i in 2:length(CN)){
    nm<-CN[i]
    res<-sapply(l,getCM,nm=nm,keepOrder=keepOrder)
    if(is.matrix(res)){
    resl[[nm]]<-as.vector(dist(t(res),method=dist))
    }
  }
  resm<-do.call(cbind,resl)
  return(resm)
}

calcCentralityExternalDistances<-function(m,l,keepOrder=FALSE,dist='euclidean'){
  CN  <- c("ID","DEG","BET","CC","SL","mnSP","PR","sdSP")
  resl<-list()
  for(i in 2:length(CN)){
    nm<-CN[i]
    rm<-getCM(m,nm=nm,keepOrder=keepOrder)
    res<-sapply(l,getCM,nm=nm,keepOrder=keepOrder)
    if(is.matrix(res)){
    cmm<-cbind(rm,res)
    cmd<-as.matrix(dist(t(cmm),method=dist))
    resl[[nm]]<-as.vector(cmd[-1,1])
    }
  }
  resm<-do.call(cbind,resl)
  return(resm)
}

evalCentralitySignificance<-function(dmi,dme){
  nmi<-colnames(dmi)
  nme<-colnames(dme)
  nms<-intersect(nmi,nme)
  l<-list()
  for(nm in nms){
    mi<-dmi[,colnames(dmi)==nm]
    me<-dme[,colnames(dme)==nm]
    ks<-ks.test(mi,me)
    l[[nm]]<-list(ks=ks,
                  dt=data.frame(val=c(mi,me),
                                cl=factor(c(rep('perm',length(mi)),
                                     rep('graph',length(me))))))
  }
  return(l)
}




```

# Prepare graph
```{r net_from_SynaptomeDB}

cid<-match('Presynaptic',getCompartments()$Name) # Let's get the ID for presynaptic compartment
cid
t<-getAllGenes4Compartment(cid) # Now we need to collect all the gene IDs for presinaptic  compartment
dim(t)
head(t)
gg<-buildFromSynaptomeByEntrez(t$HumanEntrez) # finally, build the graph using respecctive gene EntrezIDs as node IDs
summary(gg)

```

# Prepare centrality measure dataframe
```{r get.gg.cm.df}
m<-getCentralityMatrix(gg)
```

# Generate random graphs and get their centrality measure dataframes
## Erdos-Renyi model
```{r get.rgc.gnp.df}
gnp<-list()
for(i in 1:100){
  gnp[[i]]<-getRandomGraphCentrality(gg,type = 'gnp')
}
```

## Barabasi-Albert model
```{r get.rgc.pa.df}
pa<-list()
for(i in 1:100){
  pa[[i]]<-getRandomGraphCentrality(gg,type = 'pa')
}
```

## Adjacency matrix perturbation 
```{r get.rgc.cgnp.df}
cgnp<-list()
for(i in 1:100){
  cgnp[[i]]<-getRandomGraphCentrality(gg,type = 'cgnp')
}
```

## Rewiring graph
```{r get.rgc.rw.df}
rw<-list()
for(i in 1:100){
  rwg<-rewire(gg,keeping_degseq(niter = 100))
  rw[[i]]<-getCentralityMatrix(rwg)
}
```

# Calculate intra-perturbation distances for all samples with sorting
## Erdos-Renyi model
```{r get.intra.gnp.mat.ord}
gnpIDist<-calcCentralityInternalDistances(gnp)
summary(gnpIDist)
```

## Barabasi-Albert model
```{r get.intra.pa.mat.ord}
paIDist<-calcCentralityInternalDistances(pa)
summary(paIDist)
```

## Adjacency matrix perturbation 
```{r get.intra.cgnp.mat.ord}
cgnpIDist<-calcCentralityInternalDistances(cgnp)
summary(cgnpIDist)
```

## Rewiring graph
```{r get.intra.rw.mat.ord}
rwIDist<-calcCentralityInternalDistances(rw)
summary(rwIDist)
```

# Calculate distances between original and random graphs for all samples with sorting
## Erdos-Renyi model
```{r get.ex.gnp.mat.ord}
gnpEDist<-calcCentralityExternalDistances(m,gnp)
summary(gnpEDist)
```

## Barabasi-Albert model
```{r get.ex.pa.mat.ord}
paEDist<-calcCentralityExternalDistances(m,pa)
summary(paEDist)
```

## Adjacency matrix perturbation 
```{r get.ex.cgnp.mat.ord}
cgnpEDist<-calcCentralityExternalDistances(m,cgnp)
summary(cgnpEDist)
```

## Rewiring graph
```{r get.ex.rw.mat.ord}
rwEDist<-calcCentralityExternalDistances(m,rw)
summary(rwEDist)
```

# Compare distances between random graphs and random and original graphs
## Erdos-Renyi model
```{r comp.gnp.mat.ord}
gnpL<-evalCentralitySignificance(gnpIDist,gnpEDist)
gnpDF<-ldply(gnpL,function(.x){df<-.x$dt;return(df)})
names(gnpDF)[1]<-'centrality'
```

```{r plot.gnp.mat.ord,fig.height=10,fig.width=10}
ggplot(data=gnpDF,aes(y=val,x=cl,fill=cl))+geom_violin()+facet_wrap(~centrality, scales = 'free_y')
```

## Barabasi-Albert model
```{r comp.pa.mat.ord}
paL<-evalCentralitySignificance(paIDist,paEDist)
paDF<-ldply(paL,function(.x){df<-.x$dt;return(df)})
names(paDF)[1]<-'centrality'
```

```{r plot.pa.mat.ord,fig.height=10,fig.width=10}
ggplot(data=paDF,aes(y=val,x=cl,fill=cl))+geom_violin()+facet_wrap(~centrality, scales = 'free_y')
```

## Adjacency matrix perturbation 
```{r comp.cgnp.mat.ord}
cgnpL<-evalCentralitySignificance(cgnpIDist,cgnpEDist)
cgnpDF<-ldply(cgnpL,function(.x){df<-.x$dt;return(df)})
names(cgnpDF)[1]<-'centrality'
```

```{r plot.cgnp.mat.ord,fig.height=10,fig.width=10}
ggplot(data=cgnpDF,aes(y=val,x=cl,fill=cl))+geom_violin()+facet_wrap(~centrality, scales = 'free_y')
```

## Rewiring graph
```{r comp.rw.mat.ord}
rwL<-evalCentralitySignificance(rwIDist,rwEDist)
rwDF<-ldply(rwL,function(.x){df<-.x$dt;return(df)})
names(rwDF)[1]<-'centrality'
```

```{r plot.rw.mat.ord,fig.height=10,fig.width=10}
ggplot(data=rwDF,aes(y=val,x=cl,fill=cl))+geom_violin()+facet_wrap(~centrality, scales = 'free_y')
```


# Calculate intra-perturbation distances for all samples with sorting
## Erdos-Renyi model
```{r get.intra.gnp.mat.nord}
gnpIDist<-calcCentralityInternalDistances(gnp,keepOrder = TRUE)
summary(gnpIDist)
```

## Barabasi-Albert model
```{r get.intra.pa.mat.nord}
paIDist<-calcCentralityInternalDistances(pa,keepOrder = TRUE)
summary(paIDist)
```

## Adjacency matrix perturbation 
```{r get.intra.cgnp.mat.nord}
cgnpIDist<-calcCentralityInternalDistances(cgnp,keepOrder = TRUE)
summary(cgnpIDist)
```

## Rewiring graph
```{r get.intra.rw.mat.nord}
rwIDist<-calcCentralityInternalDistances(rw,keepOrder = TRUE)
summary(rwIDist)
```

# Calculate distances between original and random graphs for all samples without sorting
## Erdos-Renyi model
```{r get.ex.gnp.mat.nord}
gnpEDist<-calcCentralityExternalDistances(m,gnp,keepOrder = TRUE)
summary(gnpEDist)
```

## Barabasi-Albert model
```{r get.ex.pa.mat.nord}
paEDist<-calcCentralityExternalDistances(m,pa,keepOrder = TRUE)
summary(paEDist)
```

## Adjacency matrix perturbation 
```{r get.ex.cgnp.mat.nord}
cgnpEDist<-calcCentralityExternalDistances(m,cgnp,keepOrder = TRUE)
summary(cgnpEDist)
```

## Rewiring graph
```{r get.ex.rw.mat.nord}
rwEDist<-calcCentralityExternalDistances(m,rw,keepOrder = TRUE)
summary(rwEDist)
```

# Compare distances between random graphs and random and original graphs
## Erdos-Renyi model
```{r comp.gnp.mat.nord}
gnpL<-evalCentralitySignificance(gnpIDist,gnpEDist)
gnpDF<-ldply(gnpL,function(.x){df<-.x$dt;return(df)})
names(gnpDF)[1]<-'centrality'
```
```{r comp.gnp.ks.nord}
gnpKS<-ldply(gnpL,function(.x){df<-.x$ks$p.value;return(df)})
names(gnpKS)[1]<-'centrality'
pander(gnpKS)
```

```{r plot.gnp.mat.nord,fig.height=10,fig.width=10}
ggplot(data=gnpDF,aes(y=val,x=cl,fill=cl))+geom_violin()+facet_wrap(~centrality, scales = 'free_y')
```

## Barabasi-Albert model
```{r comp.pa.mat.nord}
paL<-evalCentralitySignificance(paIDist,paEDist)
paDF<-ldply(paL,function(.x){df<-.x$dt;return(df)})
names(paDF)[1]<-'centrality'
```
```{r comp.pa.ks.nord}
paKS<-ldply(paL,function(.x){df<-.x$ks$p.value;return(df)})
names(paKS)[1]<-'centrality'
pander(paKS)
```

```{r plot.pa.mat.nord,fig.height=10,fig.width=10}
ggplot(data=paDF,aes(y=val,x=cl,fill=cl))+geom_violin()+facet_wrap(~centrality, scales = 'free_y')
```

## Adjacency matrix perturbation 
```{r comp.cgnp.mat.nord}
cgnpL<-evalCentralitySignificance(cgnpIDist,cgnpEDist)
cgnpDF<-ldply(cgnpL,function(.x){df<-.x$dt;return(df)})
names(cgnpDF)[1]<-'centrality'
```

```{r plot.cgnp.mat.nord,fig.height=10,fig.width=10}
ggplot(data=cgnpDF,aes(y=val,x=cl,fill=cl))+geom_violin()+facet_wrap(~centrality, scales = 'free_y')
```

## Rewiring graph
```{r comp.rw.mat.nord}
rwL<-evalCentralitySignificance(rwIDist,rwEDist)
rwDF<-ldply(rwL,function(.x){df<-.x$dt;return(df)})
names(rwDF)[1]<-'centrality'
```

```{r plot.rw.mat.nord,fig.height=10,fig.width=10}
ggplot(data=rwDF,aes(y=val,x=cl,fill=cl))+geom_violin()+facet_wrap(~centrality, scales = 'free_y')
```

# Appendix {.tabset}
## Functions
```{r functions, eval=FALSE, include=TRUE}
```

## Setup R
```{r setup, eval=FALSE}
```

## Versions
### Document version
```{r docVersion, echo=FALSE, results='asis', cache=FALSE}
cat(params$version)
```

### Session Info
```{r sessionInfo, echo=FALSE, results='asis', class='text', warning=FALSE}
si<-devtools::session_info()
cat('Platform\n\n')
pander::pander(si$platform)
cat('Packages\n\n')
knitr::kable(as.data.frame(si$packages)[,c('ondiskversion','loadedversion','date','source')],align = c('l','l'))
```

